---
title: "COVID Data Analysis and Prediction"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(flexdashboard)
library(tidyverse)
library(dplyr)
library(reshape2)
library(zoo)
```


# Viewing the time series
## Column {.sidebar}
```{r,echo = FALSE, warning = FALSE}
library(flexdashboard)
dateInput("startDate", label = "Choose the start date: ", 
                value = "01-25-2020",
                min = "2020-01-25", max = as.Date(format(Sys.time())),
                format = "yyyy-mm-dd",
                startview = "month",
                weekstart = 0,
                language = "en",
                width = NULL,
                autoclose = TRUE)

dateInput("endDate", label = "Choose the end date: ",
                value = as.Date(format(Sys.time())),
                min = "2020-01-25", max = as.Date(format(Sys.time())),
                format = "yyyy-mm-dd",
                startview = "month",
                weekstart = 0,
                language = "en",
                width = NULL,
                autoclose = TRUE
          )

tempData = read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
countries = tempData[[2]]
countries = c(countries, "World")

countries = sort(countries)
selectInput("country", "Region",
  countries,
  selected = "World",
  multiple = FALSE,
  selectize = TRUE,
  width = NULL,
  size = NULL)

checkboxInput("movingAverage", "Show the 7-day moving average", value = FALSE, width = NULL)
radioButtons(
  "type",
  "Type of data",
  choices = list("Total number of cases", "Daily increase in the number of cases")
  )

radioButtons(
  "info",
  "Type of Case",
  choices = list("Confirmed", "Deaths")
  )

radioButtons(
        "predictionType",
        "Select the model you would like to use for forecasting and go to the next page to view the forecast result",
        choices = list("ARIMA", "SVM", "Polynomial Regression")
)

```


## Column (data-width = 700)
### Analyzing the time series for a desired region
```{r, echo = FALSE, warning = FALSE}
fileUrl = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_"


renderPlot({
        if (input$info == "Confirmed") filename = "confirmed_global.csv"
        #else if (input$info == "Recovered") filename = "recovered_global.csv" 
        else filename = "deaths_global.csv"
        rawData = read_csv(paste0(fileUrl, filename))

        cumulativeData = rawData[,c(-1,-3,-4)] %>% 
        group_by(`Country/Region`) %>% 
        summarise(across(everything(), sum)) %>% 
        ungroup()
        countries = cumulativeData[[1]]
        timeSeries = cumulativeData[,-1]
        dates = colnames(timeSeries)
        cumulativeData = t(timeSeries) %>% as.data.frame()
        colnames(cumulativeData) = countries
        rownames(cumulativeData) = NULL
        cumulativeData = cbind("Date" = dates, cumulativeData)
        cumulativeData$Date = as.Date(cumulativeData$Date, "%m/%d/%y")
        world = rowSums(cumulativeData[,-1])
        cumulativeData$World = as.numeric(world)
        

        # Calculating the daily cases
        dailyData = NULL
        for(i in seq(2, ncol(cumulativeData)-1)){
        dailyData = cbind(dailyData,diff(cumulativeData[[i]]))
        }
        dailyData = as.data.frame(dailyData)
        colnames(dailyData) = countries
        dailyData = cbind(Date = as.Date(dates[-1], "%m/%d/%y"), dailyData)
        dailyData$World = diff(cumulativeData$World)
        
        # Moving average for the cumulative data
        movingAverageCumulative = cbind("Date" = as.Date(dates, "%m/%d/%y"), rbind(matrix(0, 6,193),rollmean(cumulativeData[-1], k = 7)))
        movingAverageCumulative = as.data.frame(movingAverageCumulative) 
        combinedDataCumulative = cbind(movingAverageCumulative %>% select('MovingAverage' = input$country), cumulativeData %>% select('TotalCases' = input$country))
        combinedDataCumulative = cbind("Date" = as.Date(dates, "%m/%d/%y"), combinedDataCumulative)
        combinedDataCumulativeToPlot = combinedDataCumulative %>% pivot_longer(!Date, names_to = "Type", values_to = "TotalCases")
        combinedDataCumulativeToPlot = combinedDataCumulativeToPlot[combinedDataCumulativeToPlot$Date >= input$startDate & combinedDataCumulativeToPlot$Date <= input$endDate,]
        
        # Moving average for the daily cases
        movingAverageDaily = cbind("Date" = as.Date(dates[-1], "%m/%d/%y"), rbind(matrix(0, 6,193),rollmean(dailyData[-1], k = 7)))
        movingAverageDaily = as.data.frame(movingAverageDaily)
        combinedDataDaily = cbind(movingAverageDaily %>% select('MovingAverage' = input$country), dailyData %>% select('DailyIncrease' = input$country))
        combinedDataDaily = cbind("Date" = as.Date(dates[-1], "%m/%d/%y"), combinedDataDaily)
        combinedDataDailyToPlot = combinedDataDaily %>% pivot_longer(!Date, names_to = "Type", values_to = "DailyCases")
        combinedDataDailyToPlot = combinedDataDailyToPlot[combinedDataDailyToPlot$Date >= input$startDate & combinedDataDailyToPlot$Date <= input$endDate,] 
        
        if(input$type == "Total number of cases"){
                if(input$movingAverage == FALSE){
                        selectedData = cumulativeData %>% select(Date, input$country)  
                        elongatedData = selectedData %>% pivot_longer(!Date, names_to = "Country", values_to = "Count")
                        dataToPlot = elongatedData[elongatedData$Date >= input$startDate & elongatedData$Date <= input$endDate,]   
                        ggplot(dataToPlot, aes(x = Date, y = Count, col = Country)) + geom_line() + xlab("Date") + ylab(paste0("Total number of cases (", input$info, ")")) + theme_minimal() + theme(legend.position = "none")
                }else{
                        ggplot(combinedDataCumulativeToPlot, aes(x = Date, y = TotalCases, col = Type)) + geom_line() + theme(legend.title = element_blank()) + xlab("Date") + ylab(paste0("Total number of cases (", input$info, ")")) + theme_minimal() + theme(  legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6)) 
                }
        }
        else{
                if(input$movingAverage == FALSE){
                        selectedData = dailyData %>% select(Date, input$country)  
                        elongatedData = selectedData %>% pivot_longer(!Date, names_to = "Country", values_to = "Count")
                        dataToPlot = elongatedData[elongatedData$Date >= input$startDate & elongatedData$Date <= input$endDate,]   
                        ggplot(dataToPlot, aes(x = Date, y = Count, col = Country)) + geom_line()  + xlab("Date") + ylab(paste0("Daily increase in the number of cases (", input$info, ")")) + theme_minimal() + theme(legend.position = "none")
                }else{
                        ggplot(combinedDataDailyToPlot, aes(x = Date, y = DailyCases, col = Type)) + geom_line() + theme(legend.title = element_blank()) + xlab("Date") + ylab(paste0("Daily increase in the number of cases (", input$info, ")")) + theme_minimal() + theme(  legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6))
                }   
        }
})

```



# Forecast
### The plot of the forecast for the next 28 days based on the data from 01/25/2020 for the country selected in the previous page is shown here
```{r, echo = FALSE, warning = FALSE}
fileUrl = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_"
library(e1071)
library(forecast)

renderPlot({
        if (input$info == "Confirmed") filename = "confirmed_global.csv"
        else filename = "deaths_global.csv"
        rawData = read_csv(paste0(fileUrl, filename))

        cumulativeData = rawData[,c(-1,-3,-4)] %>% 
        group_by(`Country/Region`) %>% 
        summarise(across(everything(), sum)) %>% 
        ungroup()
        countries = cumulativeData[[1]]
        timeSeries = cumulativeData[,-1]
        dates = colnames(timeSeries)
        cumulativeData = t(timeSeries) %>% as.data.frame()
        colnames(cumulativeData) = countries
        rownames(cumulativeData) = NULL
        cumulativeData = cbind("Date" = dates, cumulativeData)
        cumulativeData$Date = as.Date(cumulativeData$Date, "%m/%d/%y")
        world = rowSums(cumulativeData[,-1])
        cumulativeData$World = as.numeric(world)
        

        # Calculating the daily cases
        dailyData = NULL
        for(i in seq(2, ncol(cumulativeData)-1)){
        dailyData = cbind(dailyData,diff(cumulativeData[[i]]))
        }
        dailyData = as.data.frame(dailyData)
        colnames(dailyData) = countries
        dailyData = cbind(Date = as.Date(dates[-1], "%m/%d/%y"), dailyData)
        dailyData$World = diff(cumulativeData$World)
        
        # Moving average for the cumulative data
        movingAverageCumulative = cbind("Date" = as.Date(dates, "%m/%d/%y"), rbind(matrix(0, 6,193),rollmean(cumulativeData[-1], k = 7)))
        movingAverageCumulative = as.data.frame(movingAverageCumulative) 
        combinedDataCumulative = cbind(movingAverageCumulative %>% select('MovingAverage' = input$country), cumulativeData %>% select('TotalCases' = input$country))
        combinedDataCumulative = cbind("Date" = as.Date(dates, "%m/%d/%y"), combinedDataCumulative)
        
        # Moving average for the daily cases
        movingAverageDaily = cbind("Date" = as.Date(dates[-1], "%m/%d/%y"), rbind(matrix(0, 6,193),rollmean(dailyData[-1], k = 7)))
        movingAverageDaily = as.data.frame(movingAverageDaily)
        combinedDataDaily = cbind(movingAverageDaily %>% select('MovingAverage' = input$country), dailyData %>% select('DailyIncrease' = input$country))
        combinedDataDaily = cbind("Date" = as.Date(dates[-1], "%m/%d/%y"), combinedDataDaily)
        
        if(input$type == "Total number of cases"){
                if(input$predictionType == "SVM"){
                        
                        x = as.numeric(cumulativeData$Date) - min(as.numeric(cumulativeData$Date)) + 1
                        y = combinedDataCumulative$MovingAverage
                        svmData = data.frame(x, y)
                        svmModel = svm(y~x,data = svmData, type="eps-regression", kernel ="radial")
                        predX = c(x, seq(x[length(x)]+1,x[length(x)]+28))
                        predY = predict(svmModel, newdata = data.frame(x = predX))
                        toPlot = rbind(data.frame("X" = predX, "Y" = predY, "Type" = "Prediction"), data.frame("X" = x, "Y" = y, "Type" = "Original"))
                        ggplot(toPlot, aes(X,Y, col = Type)) + geom_line() + ggtitle("Forecast using Support Vector Machine") + ylab(paste0("Total number of cases (",input$info,")")) + xlab("Days since 01/25/2020") + theme_minimal() + theme(  legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6))
                        
                }else if(input$predictionType == "Polynomial Regression"){
                        x = as.numeric(cumulativeData$Date) - min(as.numeric(cumulativeData$Date)) + 1
                        y = combinedDataCumulative$MovingAverage
                        polyData = data.frame(x, y)
                        polyModel = lm(y ~ poly(x , 4, raw=TRUE), data = polyData)
                        predX = c(x, seq(x[length(x)]+1,x[length(x)]+28))
                        predY = predict(polyModel, newdata = data.frame(x = predX))
                        toPlot = rbind(data.frame("X" = predX, "Y" = predY, "Type" = "Prediction"), data.frame("X" = x, "Y" = y, "Type" = "Original"))
                        ggplot(toPlot, aes(X,Y, col = Type)) + geom_line() + ggtitle("Forecast using polynomial regression of order 4") + ylab(paste0("Total number of cases (",input$info,")")) + xlab("Days since 01/25/2020") + theme_minimal() + theme(  legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6))
                         
                }
                else{
                        x = as.numeric(cumulativeData$Date) - min(as.numeric(cumulativeData$Date)) + 1
                        y = combinedDataCumulative$MovingAverage
                        polyData = data.frame(x, y)
                        predX = c(x, seq(x[length(x)]+1,x[length(x)]+28))
                        autoArimaFit <- auto.arima(y)
                        predY = forecast(autoArimaFit, h=28)
                        plot(predY,title("Forecast using ARIMA"))
                }
        }
        else{
                if(input$predictionType == "SVM"){
                        x = as.numeric(dailyData$Date) - min(as.numeric(dailyData$Date)) + 1
                        y = combinedDataDaily$MovingAverage
                        svmData = data.frame(x, y)
                        svmModel = svm(y~x,data = svmData, type="eps-regression", kernel ="radial")
                        predX = c(x, seq(x[length(x)]+1,x[length(x)]+28))
                        predY = predict(svmModel, newdata = data.frame(x = predX))
                        toPlot = rbind(data.frame("X" = predX, "Y" = predY, "Type" = "Prediction"), data.frame("X" = x, "Y" = y, "Type" = "Original")) 
                        ggplot(toPlot, aes(X,Y, col = Type)) + geom_line() + ggtitle("Forecast using Support Vector Machine") + ylab(paste0("Daily increase in the number of cases (",input$info,")")) + xlab("Days since 01/25/2020") + theme_minimal() + theme(  legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6))
                }else if(input$predictionType == "Polynomial Regression"){
                        x = as.numeric(dailyData$Date) - min(as.numeric(dailyData$Date)) + 1
                        y = combinedDataDaily$MovingAverage
                        polyData = data.frame(x, y)
                        polyModel = lm(y ~ poly(x , 4, raw=TRUE), data = polyData)
                        predX = c(x, seq(x[length(x)]+1,x[length(x)]+28))
                        predY = predict(polyModel, newdata = data.frame(x = predX))
                        toPlot = rbind(data.frame("X" = predX, "Y" = predY, "Type" = "Prediction"), data.frame("X" = x, "Y" = y, "Type" = "Original")) + theme_minimal() 
                        ggplot(toPlot, aes(X,Y, col = Type)) + geom_line() + ggtitle("Forecast using polynomial regression of order 4") + ylab(paste0("Daily increase in the number of cases (",input$info,")")) + xlab("Days since 01/25/2020") + theme_minimal() + theme(  legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6))
                }   
                else{
                        x = as.numeric(dailyData$Date) - min(as.numeric(dailyData$Date)) + 1
                        y = combinedDataDaily$MovingAverage
                        polyData = data.frame(x, y)
                        predX = c(x, seq(x[length(x)]+1,x[length(x)]+28))
                        autoArimaFit <- auto.arima(y)
                        predY = forecast(autoArimaFit, h=28)
                        plot(predY, ylab = paste0("Total number of cases (",input$info,")"))
                }
        }
})

```


# Further information
#### This app is inspired by the documentation, "Coronavirus (COVID-19) Visualization & Prediction". 2021. Kaggle.Com. https://www.kaggle.com/therealcyberlord/coronavirus-covid-19-visualization-prediction/log#Predictions-for-confirmed-coronavirus-cases-worldwide.

#### Source of data: "Cssegisanddata/COVID-19". 2021. Github. https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data.

#### App creator: Sreelakshmi Rajendrakumar




